
```{r}
library(dplyr)
library('gtools')
library(ggplot2)
```

```{r}
inTeams<-read.csv("https://projects.fivethirtyeight.com/soccer-api/club/spi_global_rankings.csv")
#inTeams<-read.csv('./538/spi_global_rankings.csv')
head(inTeams)
```

```{r}
mean.off<-mean(inTeams$off)
mean.def<-mean(inTeams$def)
lmean.off<-mean(log(inTeams$off))
lmean.def<-mean(log(inTeams$def) ) 
mean.off
mean.def
```

```{r}
premLeague<-subset(inTeams,league=="Barclays Premier League")
premTeams<-premLeague$name
rownames(premLeague)<-premTeams
df.prem<-premLeague[,c("off","def")]
df.prem
```

```{r}
# for poisson distribution, there's a parameter which is lambda
# when team i is playing with team j
# alpha is the offense parameter for team i
# delta is the defense parameter for team j
lmean.def<- log(mean(df.prem$def))
lmean.off<- log(mean(df.prem$off))               
df.prem["alpha"]<-log(df.prem["off"])-lmean.def
df.prem["delta"]<-lmean.off-log(df.prem["def"])
head(df.prem)
```

```{r}
alphaList<-df.prem$alpha
deltaList<-df.prem$delta
names(alphaList)<-rownames(df.prem)
names(deltaList)<-rownames(df.prem)
alphaList["Liverpool"] # for Liverpool, the alpha is 1.161783
```

```{r}
rpois(10,exp(alphaList["Liverpool"]-deltaList["Manchester City"]) )
# simulating a soccer match between Liverpool and Manchester City using the parameters alpha and delta.
```

```{r}
c(rpois(1,exp(alphaList["Liverpool"]-deltaList["Manchester City"])),
  rpois(1,exp(alphaList["Manchester City"]-deltaList["Liverpool"])))
#  from both teams' perspectives.
```

```{r}
draw.score<-function(team1,team2){
    c(
        rpois(1,exp(alphaList[team1]-deltaList[team2])),
  rpois(1,exp(alphaList[team2]-deltaList[team1]))
    )
}
draw.score("Liverpool","Arsenal")
```

```{r}
df.prem[c("Liverpool","Arsenal"),]
```

```{r}
draw.score("Liverpool","Arsenal")
```

```{r}
# All possible matches in a season
allMatches<-permutations(20, 2, v=rownames(df.prem),repeats.allowed=FALSE)
colnames(allMatches)<-c("home","away")
head(allMatches,9)
length(allMatches)
```

```{r}
# Example scores through the entire season
ScoresMatrix <- matrix(nrow=nrow(allMatches),  ncol=4)
for (ii in 1:nrow(allMatches)  ) {
     ScoresMatrix[ii,1:2]=allMatches[ii,]
     ScoresMatrix[ii,3:4]= draw.score(allMatches[ii,"home"],allMatches[ii,"away"] )  
}
colnames(ScoresMatrix)<-c("home.team","away.team","home.score","away.score")
head(ScoresMatrix)
```
```{r}
Revenue_effects <- c(149.6, 145.9, 142.1, 138.4, 73.7, 70.0, 55.2, 33.5, 29.8, 26.0,
                     22.3, 18.6, 14.9, 11.2, 7.5, 3.7, 0, -88.7, -92.5, -96.2)
League_Position <- 1:20
LeaguePositionRevenueEffect <- data.frame(League_Position = League_Position, Revenue_Effect = Revenue_effects)
```


```{r}
# Revenue effects (already defined in your code)
# LeaguePositionRevenueEffect <- data.frame(...)
# Function to simulate scores for one match
draw.score <- function(team1, team2) {
  c(
    rpois(1, exp(alphaList[team1] - deltaList[team2])),
    rpois(1, exp(alphaList[team2] - deltaList[team1]))
  )
}
```

```{r}
# Simulate an entire season
# n_sims: the number of simulations to run for the season.
simulate_season <- function(n_sims) {
  ScoresMatrix <- matrix(nrow = nrow(allMatches), ncol = 4)
  for (i in 1:n_sims) {
    tempMatches <- allMatches
    for (j in 1:nrow(tempMatches)) { # creates a temporary copy of the allMatches data frame to be used within each simulation loop.
      home.team <- tempMatches[j, "home"]
      away.team <- tempMatches[j, "away"]
      scores <- draw.score(home.team, away.team)
      ScoresMatrix[j, 1:2] <- c(home.team, away.team)
      ScoresMatrix[j, 3:4] <- scores
    }
  }
  colnames(ScoresMatrix) <- c("home.team", "away.team", "home.score", "away.score")
  return(ScoresMatrix)
}
```

```{r}
# Process results of a simulated season
process_results <- function(ScoresMatrix) {
  n_teams <- nrow(df.prem)
  team_names <- rownames(df.prem)
  points <- rep(0, n_teams)
  goals_for <- rep(0, n_teams)
  goals_against <- rep(0, n_teams)
  tiebreakers <- runif(n_teams)
  
  for (i in 1:nrow(ScoresMatrix)) {
    home.team <- ScoresMatrix[i, "home.team"]
    away.team <- ScoresMatrix[i, "away.team"]
    home.score <- as.numeric(ScoresMatrix[i, "home.score"])  # Ensure numeric type
    away.score <- as.numeric(ScoresMatrix[i, "away.score"])  # Ensure numeric type
    
    # Check data types
    if (!is.numeric(home.score) || !is.numeric(away.score)) {
      print("Error: Scores are not numeric.")
      break
    }
    
    points[team_names == home.team] <- points[team_names == home.team] + ifelse(home.score > away.score, 3, ifelse(home.score == away.score, 1, 0))
    points[team_names == away.team] <- points[team_names == away.team] + ifelse(away.score > home.score, 3, ifelse(away.score == home.score, 1, 0))
    
    goals_for[team_names == home.team] <- goals_for[team_names == home.team] + home.score
    goals_against[team_names == home.team] <- goals_against[team_names == home.team] + away.score
    goals_for[team_names == away.team] <- goals_for[team_names == away.team] + away.score
    goals_against[team_names == away.team] <- goals_against[team_names == away.team] + home.score
  }
  
  df <- data.frame(Team = team_names, Points = points, Goals.For = goals_for, Goals.Against = goals_against, Tiebreaker = tiebreakers)
  return(df)
}
```


```{r}
# Rank teams and assign prizes based on results
rank_and_assign_prizes <- function(results) {
  results <- arrange(results, desc(Points))
  results <- arrange(results, desc(Goals.For - Goals.Against), Points)
  results <- arrange(results, Tiebreaker, Points, desc(Goals.For - Goals.Against))
  
  results$League.Position <- 1:nrow(results)
  results$Prize.Money <- LeaguePositionRevenueEffect[, 2]
  return(results)
}
```

```{r}
# Define number of simulations
n_sims <- 1000
# Simulate multiple seasons
simulated_seasons <- lapply(1:n_sims, function(i) simulate_season(1))
```

############## Question 1 #################
# Assemble a ranking of the teams from best to worst according to their expected position in the table.
# Illustrate this ranking with a clear visualization of the expected position
```{r}
# Combine the results from all simulations
combined_results <- function(simulated_seasons) {
  # Make the results into a dataframe
  combined_df <- data.frame(Team = rownames(df.prem), Average_Points = numeric(nrow(df.prem)))
  # Calculate average points for each team
  for (i in 1:length(simulated_seasons)) {
    season_results <- process_results(simulated_seasons[[i]])
    combined_df$Average_Points <- combined_df$Average_Points + season_results$Points
  }
  combined_df$Average_Points <- combined_df$Average_Points / length(simulated_seasons)
  return(combined_df)
}

```

```{r}
# Rank the teams
rank_teams <- function(combined_df) {
  ranked_teams <- combined_df[order(-combined_df$Average_Points), ]
  ranked_teams$Expected_Position <- seq_len(nrow(ranked_teams))
  return(ranked_teams)
}
visualize_expected_positions <- function(ranked_teams) {
  ggplot(ranked_teams, aes(x = reorder(Team, Expected_Position), y = Expected_Position)) +
    geom_bar(stat = "identity", fill = "skyblue") +
    labs(title = "Expected Positions of Teams", x = "Team", y = "Expected Position") +
    theme_minimal() +  
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```

```{r}
simulated_seasons <- lapply(1:n_sims, function(i) simulate_season(1))
combined_results <- combined_results(simulated_seasons)
ranked_teams <- rank_teams(combined_results)
expected_position_plot <- visualize_expected_positions(ranked_teams)
print(expected_position_plot)
# Define a function to create a table from ranked teams
create_table_from_ranked_teams <- function(ranked_teams) {
  ranked_teams_table <- ranked_teams[, c("Team", "Expected_Position", "Average_Points")]
  colnames(ranked_teams_table) <- c("Team", "Expected Position", "Avg Points")
  # Print the table
  print(knitr::kable(ranked_teams_table, caption = "Ranked Teams with Expected Positions and Average Points"))
}

# Call the function to create the table
create_table_from_ranked_teams(ranked_teams)

```

############## Question 2 #################
# Which teams have the most variability in their simulated earnings?
# Again, illustrate this variability with a clear visualization.
```{r}
Revenue_effects <- c(149.6, 145.9, 142.1, 138.4, 73.7, 70.0, 55.2, 33.5, 29.8, 26.0,
                     22.3, 18.6, 14.9, 11.2, 7.5, 3.7, 0, -88.7, -92.5, -96.2)
League_Position <- 1:20
LeaguePositionRevenueEffect <- data.frame(League_Position = League_Position, Revenue_Effect = Revenue_effects)
```

```{r}
# Redefine rank_and_assign_prizes function
rank_and_assign_prizes <- function(results) {
  # Sort results by points, goals difference, and tiebreakers
  results <- results[order(-results$Points, -results$Goals.For + results$Goals.Against, results$Tiebreaker), ]
  
  # Assign league positions and prizes based on position
  results$League.Position <- seq_len(nrow(results))
  results$Prize.Money <- LeaguePositionRevenueEffect$Revenue_Effect[results$League.Position]
  
  return(results)
}

# Define function to simulate seasons and calculate revenue
simulate_and_calculate_revenue <- function(n_sims) {
  # Initialize a data frame to store revenue for each team
  team_revenue <- data.frame(Team = rownames(df.prem), Average_Revenue = numeric(nrow(df.prem)), Average_Position = numeric(nrow(df.prem)), SD_Revenue = numeric(nrow(df.prem)))
  
  # Iterate over each simulation
  for (i in 1:n_sims) {
    # Simulate a season
    season_results <- process_results(simulate_season(1))
    
    # Rank teams and assign prizes
    season_results <- rank_and_assign_prizes(season_results)
    
    # Aggregate revenue for each team
    team_revenue$Average_Revenue <- team_revenue$Average_Revenue + tapply(season_results$Prize.Money, match(team_revenue$Team, season_results$Team), sum)
    
    # Aggregate position for each team
    team_revenue$Average_Position <- team_revenue$Average_Position + match(team_revenue$Team, season_results$Team)
  }
  
  # Calculate average revenue and standard deviation of revenue for each team
  team_revenue$Average_Revenue <- team_revenue$Average_Revenue / n_sims
  team_revenue$Average_Position <- team_revenue$Average_Position / n_sims
  
  # Calculate standard deviation of revenue for each team
  for (i in 1:n_sims) {
    # Simulate a season
    season_results <- process_results(simulate_season(1))
    
    # Rank teams and assign prizes
    season_results <- rank_and_assign_prizes(season_results)
    
    # Calculate standard deviation
    team_revenue$SD_Revenue <- team_revenue$SD_Revenue + ((tapply(season_results$Prize.Money, match(team_revenue$Team, season_results$Team), sum) - team_revenue$Average_Revenue)^2)
  }
  team_revenue$SD_Revenue <- sqrt(team_revenue$SD_Revenue / n_sims)
  
  # Return results
  return(team_revenue)
}

# Run the simulation
team_revenue_results <- simulate_and_calculate_revenue(1000)

# Order the team revenue results by average position in ascending order
team_revenue_results <- team_revenue_results[order(team_revenue_results$Average_Position), ]

# Switch the values in the columns and rename them
team_revenue_results <- team_revenue_results[, c("Team", "Average_Position", "Average_Revenue", "SD_Revenue")]
colnames(team_revenue_results) <- c("Team", "Avg Position", "Avg Revenue", "Variability")

# Print the table
print("Table of Team Revenue Information:")
print(knitr::kable(team_revenue_results, caption = "Team Revenue and Position Information"))


```

############## Question 3 #################
# Which teams stand to benefit the most in monetary terms from a "lucky win" (converting one of their simulated losses to a win, holding everyone else constant)?
# Again provide a visualization of this marginal effect
```{r}
simulate_lucky_win <- function(team, opponent) {
  # Simulate a season
  season_results <- process_results(simulate_season(1))
  # Rank teams and assign prizes
  ranked_results <- rank_and_assign_prizes(season_results)
  # Calculate revenue before lucky win
  revenue_before <- ranked_results$Prize.Money[ranked_results$Team == team]
  # Simulate a match between the specified teams with a lucky win for the 'team' parameter
  lucky_win_result <- draw.score(team, opponent)
  if (lucky_win_result[1] > lucky_win_result[2]) {
    # 'team' wins
    revenue_after <- ranked_results$Prize.Money[ranked_results$Team == team] + Revenue_effects[ranked_results$League.Position[ranked_results$Team == team]]
  } else {
    # 'opponent' wins or draw
    revenue_after <- revenue_before
  }
    # Calculate the change in revenue
  revenue_change <- revenue_after - revenue_before
  return(revenue_change)
}

```

```{r}
# Define a function to calculate revenue change for each team
calculate_revenue_change <- function(team) {
  revenue_changes <- numeric(nrow(df.prem))
  for (opponent in rownames(df.prem)) {
    if (opponent != team) {
      revenue_changes[opponent] <- simulate_lucky_win(team, opponent)
    }
  }
  return(revenue_changes)
}
# Calculate revenue changes for each team
revenue_change_results <- lapply(rownames(df.prem), calculate_revenue_change)

```

```{r}
# Define a function to visualize revenue changes
visualize_revenue_changes <- function(revenue_changes) {
  team_names <- rownames(df.prem)
  team_data <- data.frame(Team = team_names, Revenue_Change = unlist(revenue_changes))
  
  ggplot(team_data, aes(x = reorder(Team, Revenue_Change), y = Revenue_Change)) +
    geom_bar(stat = "identity", fill = "skyblue") +
    labs(title = "Change in Revenue Due to Lucky Win", x = "Team", y = "Revenue Change") +
    theme_minimal() +  
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
# Visualize revenue changes
revenue_change_plot <- visualize_revenue_changes(revenue_change_results)
print(revenue_change_plot)
# Bars: Each bar represents a team, and its height represents the change in revenue. A taller bar indicates a larger potential revenue change if that team were to have a lucky win against their opponents.
```
```{r}
# Print the numeric values of revenue change results
print("Numeric values of revenue change results:")
for (i in 1:length(revenue_change_results)) {
  cat("Team:", rownames(df.prem)[i], "\n")
  cat("Revenue Change:", revenue_change_results[[i]], "\n\n")
}
# Team: Manchester City : If the value is "0", it means there would be no change in revenue if Manchester City were to win against that particular opponent.
# Non-zero values represent the change in revenue (in monetary terms) if Manchester City were to have a lucky win against that opponent. Positive values indicate an increase in revenue, while negative values indicate a decrease.

```
############## Question 4 #################
# Finally, for each team, calculate the monetary benefit to each team from either (i) an increase of 10% to their expected goals scored (holding everyone else constant), or (ii) a decreasing in their expected goals conceded by 10% (Note that this means 40 separate simulations!)
# Use this to provide a clear visualization of the benefits of investing in offense versus defense by team
```{r}


####### if add teams vector #######


# Define a function to simulate the effect of increasing expected goals for a specific team

simulate_increase_goals_team <- function(team, increase_percentage) {

  # Simulate a season

  season_results <- process_results(simulate_season(1))

  # Rank teams and assign prizes

  ranked_results <- rank_and_assign_prizes(season_results)

  

  # Find the specific team's data

  team_data <- ranked_results[ranked_results$Team == team, ]

  

  # Calculate revenue before increasing expected goals

  revenue_before <- team_data$Prize.Money

  

  # Simulate an increase in expected goals for the specified team

  # Adjust the goals for the specified team by the increase_percentage

  goals_for <- team_data$Goals.For

  goals_for_increase <- goals_for * (1 + increase_percentage)

  

  # Update the goals for the specified team

  team_data$Goals.For <- goals_for_increase

  

  # Recalculate the rankings and revenue after the adjustment

  new_ranked_results <- rank_and_assign_prizes(team_data)

  revenue_after <- new_ranked_results$Prize.Money

  

  # Calculate the change in revenue

  revenue_change <- revenue_after - revenue_before

  

  return(list(revenue_before = revenue_before, revenue_after = revenue_after, revenue_change = revenue_change))

}


# Define a function to simulate the effect of decreasing expected goals against for a specific team

simulate_decrease_goals_against_team <- function(team, decrease_percentage) {

  # Simulate a season

  season_results <- process_results(simulate_season(1))

  # Rank teams and assign prizes

  ranked_results <- rank_and_assign_prizes(season_results)

  

  # Find the specific team's data

  team_data <- ranked_results[ranked_results$Team == team, ]

  

  # Calculate revenue before decreasing expected goals against

  revenue_before <- team_data$Prize.Money

  

  # Simulate a decrease in expected goals against for the specified team

  # Adjust the goals against for all opponents of the specified team by the decrease_percentage

  opponents <- rownames(df.prem)[rownames(df.prem) != team]

  goals_against <- team_data$Goals.Against

  goals_against_decrease <- goals_against * (1 - decrease_percentage)

  

  # Update the goals against for all opponents of the specified team

  for (opponent in opponents) {

    team_data$Goals.Against[team_data$Opponent == opponent] <- goals_against_decrease

  }

  

  # Recalculate the rankings and revenue after the adjustment

  new_ranked_results <- rank_and_assign_prizes(team_data)

  revenue_after <- new_ranked_results$Prize.Money

  

  # Calculate the change in revenue

  revenue_change <- revenue_after - revenue_before

  return(list(revenue_before = revenue_before, revenue_after = revenue_after, revenue_change = revenue_change))

}


# Vector of all teams

teams <- c("Manchester City", "Liverpool", "Arsenal", "Newcastle", "Brighton and Hove Albion", "Aston Villa", "Manchester United", "Brentford", "Chelsea",

           "Crystal Palace", "Tottenham Hotspur", "West Ham United", "Fulham", "Leicester City", "Everton", "AFC Bournemouth", "Wolverhampton",

           "Leeds United", "Southampton", "Nottingham Forest")



# Lists to store results

increase_results <- list()

decrease_results <- list()


# Loop over each team and simulate changes

for (team in teams) {

  increase_results[[team]] <- simulate_increase_goals_team(team, 0.1)

  decrease_results[[team]] <- simulate_decrease_goals_against_team(team, 0.1)

}


# Dataframe for visualization

increase_df <- data.frame(

  Team = teams,

  Revenue_Change = sapply(increase_results, function(x) x$revenue_change)

)


decrease_df <- data.frame(

  Team = teams,

  Revenue_Change = sapply(decrease_results, function(x) x$revenue_change)

)


# Visualization

ggplot(increase_df, aes(x = Team, y = Revenue_Change)) +

  geom_bar(stat = "identity", fill = "skyblue") +

  ggtitle("Revenue Change with 10% Increase in Expected Goals") +

  xlab("Team") +

  ylab("Revenue Change") +

  theme(axis.text.x = element_text(angle = 45, hjust = 1))


ggplot(decrease_df, aes(x = Team, y = Revenue_Change)) +

  geom_bar(stat = "identity", fill = "lightgreen") +

  ggtitle("Revenue Change with 10% Decrease in Expected Goals Against") +

  xlab("Team") +

  ylab("Revenue Change") +

  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
library(ggplot2)

# Define color palette
colors <- c("skyblue", "lightgreen")

# Combine data frames
combined_df <- rbind(increase_df, decrease_df)
combined_df$Strategy <- rep(c("Invest in Offense", "Invest in Defense"), each = nrow(increase_df))

# Plot combined bar chart
combined_plot <- ggplot(combined_df, aes(x = Team, y = Revenue_Change, fill = Strategy)) +
  geom_bar(stat = "identity", position = "dodge") +
  ggtitle("Comparison of Revenue Changes with Different Strategies") +
  xlab("Team") +
  ylab("Revenue Change") +
  scale_fill_manual(values = colors) +
  theme(axis.text.x = element_text(angle = 65, hjust = 1))

# Display the combined plot
print(combined_plot)



```



```




